Adversarial Training of Invariant Features for Speech Recognition

### 2017-03-21 interspeech/paper.tex
Edit introduction and background sections

\title{Adversarial Training of Invariant Features for Speech Recognition}
\name{Dmitriy Serdyuk$^1$, Kartik Audhkhasi$^2$, 
      Bhuvana Ramabhadran$^2$, Phil\'emon Brakel$^1$, Samuel Thomas$^2$,
      Yoshua Bengio$^{1\dagger}$}
\address{
  $^1$MILA, Universit\'e de Montr\'eal, Canada\\
  $^2$IBM Watson, USA\\
  $^\dagger$CIFAR Fellow}
\email{author@university.edu, coauthor@company.com}

@@

\maketitle
% 
\begin{abstract}
Recent advances in domain adaptation allow us to construct networks that are
invariant to certain labeled factors. The reverse gradient algorithm uses an
adversarial learning approach to train a network to produce the desired label
as well as to deceive a second classifier that is trained on adaptation 
labels. This is tightly connected to the ideas behind generative adversarial 
networks (GANs). Therefore, some insights for training GANs are needed for the
reverse gradient's successful application. We adapt this approach to train an 
acoustic system that is invariant to undesirable factors such as the recording
environment noise type. We also investigate speaker identity invariance and 
conduct experiments on two popular datasets.
\end{abstract}
\noindent\textbf{Index Terms}: speech recognition, adversarial training, 
reverse gradient
\section{Introduction}
\label{sec:intro}
    One of the most challenging aspects of automatic speech recognition (ASR)
    is the mismatch between the training and testing acoustic conditions. 
    During testing, a system may encounter new recording conditions, 
    microphone types, speakers, accents and types of background noises. 
    Furthermore, even if the test scenarios are seen during training, there 
    can be significant variability in their statistics. Thus, it's important to 
    develop ASR systems that are invariant to unseen acoustic conditions.

    Several model and feature based adaptation methods such as Maximum 
    Likelihood Linear Regression (MLLR),  feature-based MLLR and 
    iVectors~\citep{saon2013speaker} have been proposed to handle speaker 
    variability; and Noise Adaptive Training \citep[NAT;][]{kalinli2010noise}
    and Vector Taylor Series \citep[VTS;][]{un1998speech} to handle 
    environment variability.
    % The end-to-end part has become sort of irrelevant given that the
    % experiments are actually with hybrid systems and this version of the paper will not
    % have to be sold as being end-to-end anymore.
    With the increasing success of Deep Neural 
    Network~(DNN) acoustic models for ASR 
    \citep{hinton2012deep,seide2011conversational,sainath2011making}, 
    end-to-end systems are being 
    proposed~\citep{miao2015eesen,sainath2015learning} for modeling the 
    acoustic conditions within a single network. This allows us to take 
    advantage of the network's ability to learn highly non-linear feature 
    transformations, with greater flexibility in constructing training 
    objective functions that promote learning of noise invariant 
    representations.
    % Phil: The part below should include speaker invariance and mention the WSJ task
    The main idea of this work is to force the acoustic model 
    to learn a representation which is invariant to noise conditions, instead of 
    explicitly using noise robust acoustic features 
    (Section~\ref{sec:invariant-speech}). This type of noise-invariant 
    training requires noise-condition labels during training only. It is 
    related to the idea of generative adversarial networks (GAN) and the 
    gradient reverse method proposed in~\cite{goodfellow2014generative} 
    and~\cite{ganin2014unsupervised} respectively 
    (Section~\ref{sec:relatedwork}). We present results on the Aurora-4 speech 
    recognition task in Section~\ref{sec:experiments} and summarize 
    our findings in Section~\ref{sec:discussion}.

% This is more of an explanation of the method rather than a literature review.
    % Perhaps we should change the section title.
\section{Related Work}
\label{sec:relatedwork}
\emph{Generative Adversarial Networks} consist of two networks: the generator and the discriminator. 
    The generator network $G$ has an
    input of randomly-generated feature vectors and is asked to produce a
    sample, e.g.\ an image, similar to the images in the training set. The discriminator network $D$
    can either receive a generated image from the generator $G$ or an image
    from the training set. Its task is to distinguish
    between the ``fake'' generated image and the ``real'' image taken from the dataset. Thus,
    the discriminator is just a classifier network with a sigmoid output layer
    and can be trained with gradient backpropagation. This gradient can be propagated further
    to the generator network (assuming that the output of the generator is
    continuous).

    The two networks in the GAN setup are competing with each other: the 
    generator is trying to deceive the discriminator network, while the discriminator tries
    to do its best to recognize if there was a deception, similar to adversarial game-theoretic settings.    
    Formally, the objective function of GAN training is
    \begin{align*}
        \min_G \max_D V(D, G) = &\mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + \\
            &\mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (1 - D(G(\bm{z})))].
    \end{align*}
    The maximization over the discriminator $D$ forms a usual cross-entropy objective, the gradients are
    computed with respect to the parameters of $D$. The parameters of $G$ are
    minimizing the classification objective using the gradients
    propagated through the second term. The minimization over $G$ makes it
    produce samples which $D$ classifies as originating from the train data.
    Several practical guidelines were proposed for optimizing GANs in~\cite{radford2015unsupervised} and 
    further explored in~\cite{salimans2016improved}.

@@

    The effectiveness of representations produced by a neural network for internal 
    noise reduction is discussed in~\cite{yu2013feature} and this work sets a 
    baseline for experiments on the Aurora-4 dataset. Recently, 
    in~\cite{yusuke2016adversarial} a multilayer sigmoidal network was trained 
    in an adversarial fashion on an in-house transcription task corrupted by noise.
\begin{figure}[t]
    \centering
    \includegraphics[width=0.4\linewidth]{model.pdf}
    \caption{The model consists of three neural networks. The encoder $E$ produces
    the intermediate representation $h$ which used in the recognizer $R$ and 
    in the domain discriminator $D$. The hidden representation $h$ is trained to improve
    the recognition and minimize the domain discriminator accuracy. The domain discriminator
    is a classifier trained to maximize its accuracy on the noise type
    classification task.}
    \label{fig:model}
\end{figure}[t]

>>>>

\title{Adversarial Training of Invariant Features for Speech Recognition}
\name{Dmitriy Serdyuk$^1$, Kartik Audhkhasi$^2$, 
      Bhuvana Ramabhadran$^2$, Phil\'emon Brakel$^1$, Samuel Thomas$^2$,
      Yoshua Bengio$^{1\dagger}$}
\address{
  $^1$MILA, Universit\'e de Montr\'eal, Canada\\
  $^2$IBM Watson, USA\\
  $^\dagger$CIFAR Fellow}
\email{serdyuk@iro.umontreal.ca}

@@

\maketitle
% 
\begin{abstract}
Recent advances in domain adaptation allow us to construct networks that are
invariant to certain labeled factors. The reverse gradient algorithm uses an
adversarial learning approach to train a network to produce the desired label
as well as to deceive a second classifier that is trained on adaptation 
labels. This is tightly connected to the ideas behind generative adversarial 
networks (GANs). Therefore, some insights for training GANs are needed for the
reverse gradient's successful application. We adapt this approach to train an 
acoustic system that is invariant to undesirable factors such as the recording
environment noise type. We conduct experiments on a popular dataset.
\end{abstract}
\noindent\textbf{Index Terms}: speech recognition, adversarial training, 
reverse gradient
\section{Introduction}
\label{sec:intro}
    One of the most challenging aspects of automatic speech recognition (ASR)
    is the mismatch between the training and testing acoustic conditions. 
    During testing, a system may encounter new recording conditions, 
    microphone types, speakers, accents and types of background noises. 
    Furthermore, even if the test scenarios are seen during training, there 
    can be significant variability in their statistics. Thus, it's important to 
    develop ASR systems that are invariant to unseen acoustic conditions.

    Several model and feature based adaptation methods such as Maximum 
    Likelihood Linear Regression (MLLR),  feature-based MLLR and 
    iVectors~\citep{saon2013speaker} have been proposed to handle speaker 
    variability; and Noise Adaptive Training \citep[NAT;][]{kalinli2010noise}
    and Vector Taylor Series \citep[VTS;][]{un1998speech} to handle 
    environment variability.
    % The end-to-end part has become sort of irrelevant given that the
    % experiments are actually with hybrid systems and this version of the paper will not
    % have to be sold as being end-to-end anymore.
    With the increasing success of Deep Neural 
    Network~(DNN) acoustic models for ASR investigated 
    in~\citep{hinton2012deep,seide2011conversational,sainath2011making}, 
    and in more recent works~\citep{miao2015eesen,sainath2015learning} 
    complicated structure of acoustic conditions is modeled within a single network.
    This allows us to take 
    advantage of the network's ability to learn highly non-linear feature 
    transformations, with greater flexibility in constructing training 
    objective functions that promote learning of noise invariant 
    representations.
    % Phil: The part below should include speaker invariance and mention the WSJ task
    The main idea of this work is to force the acoustic model 
    to learn a representation which is invariant to noise conditions, instead of 
    explicitly using noise robust acoustic features 
    (Section~\ref{sec:invariant-speech}). This type of noise-invariant 
    training requires noise-condition labels during training only. It is 
    related to the idea of generative adversarial networks (GAN) and the 
    gradient reverse method proposed in~\cite{goodfellow2014generative} 
    and~\cite{ganin2014unsupervised} respectively 
    (Section~\ref{sec:background}). Then we discuss previous work in 
    Section~\ref{sec:relatedwork}. 
    We present results on the Aurora-4 speech 
    recognition task in Section~\ref{sec:experiments} and summarize 
    our findings in Section~\ref{sec:discussion}.

% This is more of an explanation of the method rather than a literature review.
    % Perhaps we should change the section title.
\section{Background}
\label{sec:background}
\emph{Generative Adversarial Networks} consist of two networks: the generator and the discriminator. 
    The generator network $G$ has an
    input of randomly-generated feature vectors and is asked to produce a
    sample, e.g.\ an image, similar to the images in the training set. The discriminator network $D$
    can either receive a generated image from the generator $G$ or an image
    from the training set. Its task is to distinguish
    between the ``counterfeit'' generated image and the ``genuine'' image taken from the dataset. Thus,
    the discriminator is just a classifier network with a sigmoid output layer
    and can be trained with gradient backpropagation. This gradient can be propagated further
    to the generator network (assuming that the output of the generator is
    continuous).

    The two networks in the GAN setup are competing with each other: the 
    generator is trying to deceive the discriminator network, while the discriminator tries
    to do its best to recognize if there was a deception, similar to adversarial game-theoretic settings.    
    Formally, the objective function of GAN training is
    \begin{align}
        \min_G \max_D V(D, G) = &\mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + \\
            &\mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (1 - D(G(\bm{z})))].
        \label{eq:gan}
    \end{align}
    The maximization over the discriminator $D$ forms a usual cross-entropy objective, the gradients are
    computed with respect to the parameters of $D$. An important property of
    this objective is that the gradient of the composition~$D(G(\cdot))$ is well
    defined. Therefore the generator~$G$ can be trained with the backpropagation
    algorithm using the gradient signal from the discriminator~$D$.
    The generator~$G$ is
    minimizing the classification objective using the gradients
    propagated through the second term. The minimization over $G$ makes it
    produce samples which $D$ classifies as originating from the train data.

    Several practical guidelines were proposed for optimizing GANs in~\cite{radford2015unsupervised} and 
    further explored in~\cite{salimans2016improved}. The second term of Eq.~\ref{eq:gan}
    is frequently exchanged with
    \begin{equation}
        - \mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (D(G(\bm{z})))].
    \end{equation}
    The gradient of this term has the same sign but it has better properties.
    We experimented with severall losses for our discriminator and chose one
    similar to Eq.~\ref{eq:gan} but we maintain the negative sign in our notation
    to demonstrate the general direction of the gradient of the discriminator loss.

@@

    The effectiveness of representations produced by a neural network for internal 
    noise reduction is discussed in~\cite{yu2013feature} and this work sets a 
    baseline for experiments on the Aurora-4 dataset. Recently, 
    in~\cite{yusuke2016adversarial} a multilayer sigmoidal network was trained 
    in an adversarial fashion on an in-house transcription task corrupted by noise.
\begin{figure}[t]
    \centering
    \includegraphics[width=0.4\linewidth]{model.pdf}
    \caption{The model consists of three neural networks. The encoder $E$ produces
    the intermediate representation $h$ which used in the recognizer $R$ and 
    in the domain discriminator $D$. The hidden representation $h$ is trained to improve
    the recognition and minimize the domain discriminator accuracy. The domain discriminator
    is a classifier trained to maximize its accuracy on the noise type
    classification task.}
    \label{fig:model}
\end{figure}

\section{Related Work}
\label{sec:relatedwork}

%TODO
